---
title: "a_02_CalculateHeatIndex"
author: "Aline Maybank"
date: "2026-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting Up
The purpose of this script is to calculate the Heat Index for Baltimore and DC daily for the study period. The method used to calculate the heat index was created by: https://ehp.niehs.nih.gov/doi/10.1289/ehp.1206273; and found https://cran.r-project.org/web/packages/weathermetrics/index.html

## Load Packages
```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(riem)
library(weathermetrics)
library(MASS)
library(emmeans)
library(broom)
library(readr)
library(knitr)
library(kableExtra)
library(splines)
library(sf)
library(tigris)
library(patchwork)
library(cowplot)
library(stringr)
```

## Load Crime Data
```{r}
# Baltimore crime data from 2020-2024
Baltimore_allcrimes_2020_clean_flat <- read.csv("Crime_Data/2025-10-07_Baltimore_allcrimes_2020_clean_flat.csv")
Baltimore_allcrimes_2021_clean_flat <- read.csv("Crime_Data/2025-10-07_Baltimore_allcrimes_2021_clean_flat.csv")
Baltimore_allcrimes_2022_clean_flat <- read.csv("Crime_Data/2025-10-07_Baltimore_allcrimes_2022_clean_flat.csv")
Baltimore_allcrimes_2023_clean_flat <- read.csv("Crime_Data/2025-10-07_Baltimore_allcrimes_2023_clean_flat.csv")
Baltimore_allcrimes_2024_clean_flat <- read.csv("Crime_Data/2025-10-07_Baltimore_allcrimes_2024_clean_flat.csv")

# DC crime data from 2020-2024
DC_allcrimes_2020_clean_flat <- read.csv("Crime_Data/2025-10-07_DC_allcrimes_2020_clean_flat.csv")
DC_allcrimes_2021_clean_flat <- read.csv("Crime_Data/2025-10-07_DC_allcrimes_2021_clean_flat.csv")
DC_allcrimes_2022_clean_flat <- read.csv("Crime_Data/2025-10-07_DC_allcrimes_2022_clean_flat.csv")
DC_allcrimes_2023_clean_flat <- read.csv("Crime_Data/2025-10-07_DC_allcrimes_2023_clean_flat.csv")
DC_allcrimes_2024_clean_flat <- read.csv("Crime_Data/2025-10-07_DC_allcrimes_2024_clean_flat.csv")
```

# Part 1: Heat Index Calculation

## Retrieve Heat Index Data

Retrieve hourly ASOS data (2020–2024)
```{r}
date_start <- "2020-01-01"
date_end   <- "2025-01-01"
tz_local   <- "America/New_York"

# Fetch hourly ASOS
# DC: station = KDCA (Reagan National)
# Baltimore: station = KBWI (BWI)
dc_hr_raw <- riem_measures(station = "KDCA", date_start = date_start, date_end = date_end)
bal_hr_raw <- riem_measures(station = "KBWI", date_start = date_start, date_end = date_end)
```

## Clean ASOS Data
```{r}
# Function to clean and recover relative humidity
clean_asos <- function(df, tz_local) {
  df %>%
    dplyr::transmute(
      datetime_local = lubridate::with_tz(as.POSIXct(.data$valid, tz = "UTC"), tzone = tz_local),
      date  = as.Date(datetime_local),
      tmpf  = suppressWarnings(as.numeric(.data$tmpf)),
      relh  = suppressWarnings(as.numeric(.data$relh)),
      dwpf  = suppressWarnings(as.numeric(.data$dwpf))
    ) %>%
    dplyr::mutate(
      rh_from_dp = dplyr::if_else(
        is.na(relh) & !is.na(tmpf) & !is.na(dwpf),
        weathermetrics::dewpoint.to.humidity(
          dp = dwpf, t = tmpf, temperature.metric = "fahrenheit"
        ),
        NA_real_
      ),
      relh = dplyr::coalesce(relh, rh_from_dp),
      relh = pmax(pmin(relh, 100), 0),
      tmpf = dplyr::if_else(tmpf < -80 | tmpf > 150, NA_real_, tmpf)
    ) %>%
    dplyr::select(datetime_local, date, tmpf, relh)
}

dc_hr  <- clean_asos(dc_hr_raw,  tz_local)
bal_hr <- clean_asos(bal_hr_raw, tz_local)

```

## Compute Daily Heat Index
```{r}
# Vectorized Heat Index calculator
hi_vec <- function(temp_f, rh_pct) {
  suppressWarnings(weathermetrics::heat.index(
    t = temp_f,
    rh = rh_pct,
    temperature.metric = "fahrenheit",
    output.metric      = "fahrenheit",
    round = 1
  ))
}

# DC daily heat index (from hourly)
dc_daily_hi <- dc_hr %>%
  dplyr::mutate(hi_f = hi_vec(tmpf, relh)) %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(
    hi_mean_f = if (all(is.na(hi_f))) NA_real_ else mean(hi_f, na.rm = TRUE),
    hi_max_f  = if (all(is.na(hi_f))) NA_real_ else max(hi_f,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(date <= as.Date("2024-12-31")) %>%  # ADD THIS LINE
  dplyr::mutate(city = "DC")

# Baltimore daily heat index (from hourly)
bal_daily_hi <- bal_hr %>%
  dplyr::mutate(hi_f = hi_vec(tmpf, relh)) %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(
    hi_mean_f = if (all(is.na(hi_f))) NA_real_ else mean(hi_f, na.rm = TRUE),
    hi_max_f  = if (all(is.na(hi_f))) NA_real_ else max(hi_f,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(date <= as.Date("2024-12-31")) %>%  # ADD THIS LINE
  dplyr::mutate(city = "Baltimore")

# Combine for analysis
daily_hi_all <- dplyr::bind_rows(dc_daily_hi, bal_daily_hi) %>%
  dplyr::select(city, date, hi_mean_f, hi_max_f)

```

## Verify Heat Index Calculation
```{r}
summary(dc_daily_hi$hi_max_f)
summary(bal_daily_hi$hi_max_f)

# Coverage check
seq_dates <- seq(as.Date("2020-01-01"), as.Date("2024-12-31"), by = "day")
cat("Missing DC dates:", sum(!(seq_dates %in% dc_daily_hi$date)), "\n")
cat("Missing Baltimore dates:", sum(!(seq_dates %in% bal_daily_hi$date)), "\n")
```
For future reference, 2024 was a leap year.

## Export Heat Index Data
```{r}
out_dir <- "outputs/heat_index"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Export various formats
readr::write_csv(daily_hi_all, file.path(out_dir, "daily_hi_wide_f.csv"))
readr::write_csv(
  dplyr::filter(daily_hi_all, city == "DC"),
  file.path(out_dir, "daily_hi_DC_f.csv")
)
readr::write_csv(
  dplyr::filter(daily_hi_all, city == "Baltimore"),
  file.path(out_dir, "daily_hi_Baltimore_f.csv")
)
```

## Calculate HI Percentile Thresholds
```{r}
# Calculate 85th and 95th percentile thresholds by city
hi_thresholds_hw <- daily_hi_all %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    p85 = quantile(hi_max_f, 0.85, na.rm = TRUE),
    p95 = quantile(hi_max_f, 0.95, na.rm = TRUE),
    .groups = "drop"
  )

hi_thresholds_hw %>%
  dplyr::mutate(
    p85 = round(p85, 1),
    p95 = round(p95, 1)
  ) %>%
  knitr::kable(
    format = "html",
    col.names = c("City", "85th Percentile (Max HI °F)", "95th Percentile (Max HI °F)"),
    caption = "Heat Index Thresholds (Daily Max HI)"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )
```

## Visualize Heat Index Distribution
```{r}
# Percentile thresholds for plotting (by city)
hi_thresh <- daily_hi_all %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    p85_max  = stats::quantile(hi_max_f,  0.85, na.rm = TRUE),
    p95_max  = stats::quantile(hi_max_f,  0.95, na.rm = TRUE),
    p85_mean = stats::quantile(hi_mean_f, 0.85, na.rm = TRUE),
    p95_mean = stats::quantile(hi_mean_f, 0.95, na.rm = TRUE),
    .groups = "drop"
  )

# Long format for plotting (keep all cities; drop NA hi_f)
hi_long <- daily_hi_all %>%
  tidyr::pivot_longer(
    cols = c(hi_max_f, hi_mean_f),
    names_to = "metric",
    values_to = "hi_f"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(metric,
      hi_max_f  = "Daily Max HI (°F)",
      hi_mean_f = "Daily Mean HI (°F)"
    )
  ) %>%
  dplyr::filter(!is.na(hi_f))

# Plot function
plot_hi_density <- function(df_metric_label) {

  th_cols <- if (df_metric_label == "Daily Max HI (°F)") {
    c("p85_max", "p95_max")
  } else {
    c("p85_mean", "p95_mean")
  }

  tdf <- hi_thresh %>%
    dplyr::select(
      city,
      p85 = dplyr::all_of(th_cols[1]),
      p95 = dplyr::all_of(th_cols[2])
    )

  dat <- hi_long %>%
    dplyr::filter(metric == df_metric_label) %>%
    dplyr::left_join(tdf, by = "city")

  x_min <- floor(min(dat$hi_f, na.rm = TRUE) / 5) * 5
  x_max <- ceiling(max(dat$hi_f, na.rm = TRUE) / 5) * 5
  breaks_5 <- seq(x_min, x_max, by = 5)

  ggplot2::ggplot(dat, ggplot2::aes(x = hi_f)) +
    ggplot2::geom_histogram(ggplot2::aes(y = ggplot2::after_stat(density)),
                            bins = 40, alpha = 0.25) +
    ggplot2::geom_density(linewidth = 1) +
    ggplot2::geom_vline(ggplot2::aes(xintercept = p85), linetype = 2) +
    ggplot2::geom_vline(ggplot2::aes(xintercept = p95), linetype = 2) +
    ggplot2::facet_wrap(~ city, ncol = 1, scales = "free_y") +
    ggplot2::scale_x_continuous(
      breaks = breaks_5,
      labels = function(x) sprintf("%.0f°", x)
    ) +
    ggplot2::labs(
      title = paste0("Distribution of ", df_metric_label, " with 85th/95th Percentiles"),
      x = df_metric_label, y = "Density"
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(panel.grid = ggplot2::element_blank())
}

print(plot_hi_density("Daily Max HI (°F)"))
print(plot_hi_density("Daily Mean HI (°F)"))
```

# Part 2: Crime Data Preparation

## Normalize and Combine Crime Data
```{r}
# Normalization function
normalize_for_bind <- function(df, date_col, city_name, offense_col) {
  df %>%
    dplyr::mutate(
      city       = city_name,
      crime_date = as.Date(.data[[date_col]]),
      offense_cat = as.character(.data[[offense_col]]) |> trimws()
    ) %>%
    # drop conflicting mixed-type column if it exists
    dplyr::select(-dplyr::any_of("Post"))
}

# Combine all crime data
crime_all <- dplyr::bind_rows(
  normalize_for_bind(DC_allcrimes_2020_clean_flat, "START_DATE", "DC", "OFFENSE"),
  normalize_for_bind(DC_allcrimes_2021_clean_flat, "START_DATE", "DC", "OFFENSE"),
  normalize_for_bind(DC_allcrimes_2022_clean_flat, "START_DATE", "DC", "OFFENSE"),
  normalize_for_bind(DC_allcrimes_2023_clean_flat, "START_DATE", "DC", "OFFENSE"),
  normalize_for_bind(DC_allcrimes_2024_clean_flat, "START_DATE", "DC", "OFFENSE"),
  normalize_for_bind(Baltimore_allcrimes_2020_clean_flat, "CrimeDateT", "Baltimore", "Descriptio"),
  normalize_for_bind(Baltimore_allcrimes_2021_clean_flat, "CrimeDateT", "Baltimore", "Descriptio"),
  normalize_for_bind(Baltimore_allcrimes_2022_clean_flat, "CrimeDateT", "Baltimore", "Descriptio"),
  normalize_for_bind(Baltimore_allcrimes_2023_clean_flat, "CrimeDateT", "Baltimore", "Descriptio"),
  normalize_for_bind(Baltimore_allcrimes_2024_clean_flat, "CrimeDateT", "Baltimore", "Descriptio")
)

# Define violent categories
violent_baltimore <- c(
  "AGG. ASSAULT",
  "HOMICIDE",
  "RAPE",
  "ROBBERY",
  "ROBBERY - CARJACKING",
  "ROBBERY - COMMERCIAL",
  "SHOOTING"
)

violent_dc <- c(
  "ASSAULT W/DANGEROUS WEAPON",
  "HOMICIDE",
  "ROBBERY",
  "SEX ABUSE"
)

crime_all <- crime_all %>%
  dplyr::mutate(
    offense_std = stringr::str_to_upper(stringr::str_squish(as.character(offense_cat))),
    crime_type = dplyr::case_when(
      city == "Baltimore" & offense_std %in% violent_baltimore ~ "VIOLENT",
      city == "Baltimore"                                     ~ "NONVIOLENT",
      city == "DC"        & offense_std %in% violent_dc        ~ "VIOLENT",
      city == "DC"                                             ~ "NONVIOLENT",
      TRUE                                                     ~ NA_character_
    )
  )
```


## Create Daily Crime Counts
```{r}
# Build daily panel (crime counts + HI join + HI categories)
crime_daily <- crime_all %>%
  dplyr::mutate(
    dow   = lubridate::wday(crime_date, label = TRUE, abbr = TRUE),
    month = lubridate::month(crime_date, label = TRUE, abbr = TRUE),
    year  = lubridate::year(crime_date)
  ) %>%
  dplyr::group_by(city, crime_date, dow, month, year) %>%
  dplyr::summarise(
    n_total      = dplyr::n(),
    n_violent    = sum(toupper(as.character(crime_type)) == "VIOLENT", na.rm = TRUE),
    n_nonviolent = n_total - n_violent,
    .groups = "drop"
  ) %>%
  dplyr::left_join(daily_hi_all, by = c("city", "crime_date" = "date")) %>%
  dplyr::left_join(hi_thresholds_hw, by = "city") %>%
  dplyr::mutate(
    hi_cat = dplyr::case_when(
  is.na(hi_max_f)                     ~ "Missing",
  hi_max_f <  p85                     ~ "<85th",
  hi_max_f >= p85 & hi_max_f < p95    ~ "Hot (85–95th)",
  hi_max_f >= p95                     ~ "Extreme (≥95th)"
),
    hi_cat = factor(hi_cat, levels = c("<85th","Hot (85–95th)","Extreme (≥95th)","Missing"))
  )

cat("Missing HI values:", sum(is.na(crime_daily$hi_max_f)), "\n")
cat("Date range:", min(crime_daily$crime_date, na.rm = TRUE), "to",
    max(crime_daily$crime_date, na.rm = TRUE), "\n")
cat("Cities:", paste(unique(crime_daily$city), collapse = ", "), "\n")
```

Split the dataset in two (one for each city)
```{r}
crime_daily_dc <- crime_daily %>%
  dplyr::filter(city == "DC")

crime_daily_bal <- crime_daily %>%
  dplyr::filter(city == "Baltimore")
```


# Part 3: Statistical Models

## Check Overdispersion
```{r}
check_od_city <- crime_daily %>%
  dplyr::filter(hi_cat != "Missing") %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    mean_total = mean(n_total, na.rm = TRUE),
    var_total  = var(n_total,  na.rm = TRUE),
    ratio      = var_total / mean_total,
    .groups = "drop"
  )

# Print the table as a kable
check_od_city %>%
  dplyr::mutate(
    mean_total = round(mean_total, 2),
    var_total  = round(var_total, 2),
    ratio      = round(ratio, 2)
  ) %>%
  knitr::kable(
    format = "html",
    col.names = c("City", "Mean (Total Daily Crimes)", "Variance (Total Daily Crimes)", "Variance / Mean"),
    caption = "Overdispersion Check by City (Daily Total Crime Counts)"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )
```


## Fit Negative Binomial Models
```{r}
## Fit Negative Binomial Models (Total, Violent, Non-Violent)
dc_data  <- crime_daily_dc  %>% dplyr::filter(hi_cat != "Missing")
bal_data <- crime_daily_bal %>% dplyr::filter(hi_cat != "Missing")

# DC
mod_dc_total      <- MASS::glm.nb(n_total      ~ hi_cat + dow + factor(month), data = dc_data)
mod_dc_violent    <- MASS::glm.nb(n_violent    ~ hi_cat + dow + factor(month), data = dc_data)
mod_dc_nonviolent <- MASS::glm.nb(n_nonviolent ~ hi_cat + dow + factor(month), data = dc_data)

# Baltimore
mod_bal_total      <- MASS::glm.nb(n_total      ~ hi_cat + dow + factor(month), data = bal_data)
mod_bal_violent    <- MASS::glm.nb(n_violent    ~ hi_cat + dow + factor(month), data = bal_data)
mod_bal_nonviolent <- MASS::glm.nb(n_nonviolent ~ hi_cat + dow + factor(month), data = bal_data)
```


## IRR Visualization Function
```{r}
## IRR helper: returns a tidy dataframe (not a plot)
irr_df <- function(model, outcome_label, city_label) {

  em <- emmeans::emmeans(model, ~ hi_cat, type = "response")
  ct <- emmeans::contrast(em, method = "trt.vs.ctrl", ref = 1, type = "response")

  ct_df <- as.data.frame(stats::confint(ct))

  # Harmonize CI column names
  nm <- names(ct_df)
  if (!"lower.CL" %in% nm) {
    if ("asymp.LCL" %in% nm) names(ct_df)[nm == "asymp.LCL"] <- "lower.CL"
    else if ("LCL" %in% nm)  names(ct_df)[nm == "LCL"]       <- "lower.CL"
  }
  nm <- names(ct_df)
  if (!"upper.CL" %in% nm) {
    if ("asymp.UCL" %in% nm) names(ct_df)[nm == "asymp.UCL"] <- "upper.CL"
    else if ("UCL" %in% nm)  names(ct_df)[nm == "UCL"]       <- "upper.CL"
  }

  # IRR column (usually "ratio")
  ycol <- if ("ratio" %in% names(ct_df)) "ratio" else {
    cand <- intersect(c("response", "estimate", "emmean"), names(ct_df))
    if (length(cand) == 0) stop("Couldn't find IRR/estimate column in ct_df.")
    cand[1]
  }

  ct_df %>%
    dplyr::mutate(
      hi_level = sub(" / <85th", "", contrast),
      irr      = .data[[ycol]],
      outcome  = outcome_label,
      city     = city_label
    ) %>%
    dplyr::select(city, outcome, hi_level, irr, lower.CL, upper.CL)
}

```


## Generate IRR Plots
```{r}

# Build IRR tables (DC + Baltimore; Total + Violent + Non-Violent)
irr_dc <- dplyr::bind_rows(
  irr_df(mod_dc_total,      "Total",       "DC"),
  irr_df(mod_dc_violent,    "Violent",     "DC"),
  irr_df(mod_dc_nonviolent, "Non-Violent", "DC")
)

irr_bal <- dplyr::bind_rows(
  irr_df(mod_bal_total,      "Total",       "Baltimore"),
  irr_df(mod_bal_violent,    "Violent",     "Baltimore"),
  irr_df(mod_bal_nonviolent, "Non-Violent", "Baltimore")
)

plot_irr_city <- function(df_city, title) {

  df_city <- df_city %>%
    dplyr::mutate(
      # Standardize dashes + whitespace so labels match
      hi_level = stringr::str_squish(hi_level),
      hi_level = stringr::str_replace_all(hi_level, "\u2013|\u2014", "-"),  # en/em dash -> hyphen

      # Map anything containing "Hot" or "Extreme" to the exact labels you want
      hi_level = dplyr::case_when(
        stringr::str_detect(hi_level, "Hot")     ~ "Hot (85–95th)",
        stringr::str_detect(hi_level, "Extreme") ~ "Extreme (≥95th)",
        TRUE ~ hi_level
      ),

      # Force order
      hi_level = factor(hi_level, levels = c("Hot (85–95th)", "Extreme (≥95th)"))
    )

  ggplot2::ggplot(
    df_city,
    ggplot2::aes(
      x = hi_level,
      y = irr,
      ymin = lower.CL,
      ymax = upper.CL,
      shape = outcome,
      linetype = outcome
    )
  ) +
    ggplot2::geom_hline(yintercept = 1, linetype = 2) +
    ggplot2::geom_pointrange(
      position = ggplot2::position_dodge(width = 0.55)
    ) +
    ggplot2::scale_y_log10() +
    ggplot2::labs(
      title = title,
      x = "Heat-index category vs <85th",
      y = "Incidence Rate Ratio (log scale)"
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(panel.grid = ggplot2::element_blank())
}

p_irr_dc  <- plot_irr_city(irr_dc,  "DC: IRR (Total / Violent / Non-violent) vs <85th")
p_irr_bal <- plot_irr_city(irr_bal, "Baltimore: IRR (Total / Violent / Non-violent) vs <85th")

print(p_irr_dc)
print(p_irr_bal)
```


# Part 4: Publication-Ready Figures

## Combine IRR tables
```{r}
# Combine IRR tables into one plotting dataframe
irr_all_three <- dplyr::bind_rows(irr_dc, irr_bal) %>%
  dplyr::mutate(
    # Force the order you want on the x-axis
    hi_level = factor(hi_level, levels = c("Hot (85–95th)", "Extreme (≥95th)")),
    # Force legend / facet ordering
    outcome  = factor(outcome, levels = c("Total", "Violent", "Non-Violent")),
    city     = factor(city, levels = c("DC", "Baltimore"))
  ) %>%
  # rename to match your figure code below (IRR/LCL/UCL)
  dplyr::rename(
    IRR = irr,
    LCL = lower.CL,
    UCL = upper.CL
  )

# For the 2-outcome plot (Violent vs Non-Violent)
irr_combined <- irr_all_three %>%
  dplyr::filter(outcome %in% c("Violent", "Non-Violent")) %>%
  dplyr::mutate(outcome = factor(outcome, levels = c("Violent", "Non-Violent")))

```

Set y-axis limit
```{r}
y_min     <- min(irr_combined$LCL, na.rm = TRUE) * 0.95
y_max     <- max(irr_combined$UCL, na.rm = TRUE) * 1.05

y_min_all <- min(irr_all_three$LCL, na.rm = TRUE) * 0.95
y_max_all <- max(irr_all_three$UCL, na.rm = TRUE) * 1.05
```

Plot 1: Violent vs Non-Violent (faceted by outcome x city)
```{r}

p_vio_vs_nonvio <- ggplot2::ggplot(
  irr_combined,
  ggplot2::aes(x = hi_level, y = IRR, ymin = LCL, ymax = UCL)
) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2) +
  ggplot2::geom_pointrange(size = 0.8, fatten = 4) +
  ggplot2::facet_grid(outcome ~ city) +
  ggplot2::scale_y_log10(limits = c(y_min, y_max)) +
  ggplot2::labs(
    title = "Incidence Rate Ratios: Violent vs Non-Violent Crime on Hot/Extreme Days",
    y = "Incidence Rate Ratio (log scale)",
    x = "Heat Index Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text = ggplot2::element_text(size = 12, face = "bold"),
    axis.text.x = ggplot2::element_text(angle = 20, hjust = 1)
  )

print(p_vio_vs_nonvio)
```

Plot 2: Total + Violent + Non-Violent (one panel per city)
```{r}
p_all_three <- ggplot2::ggplot(
  irr_all_three,
  ggplot2::aes(
    x = hi_level, y = IRR, ymin = LCL, ymax = UCL,
    shape = outcome, linetype = outcome
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2) +
  ggplot2::geom_pointrange(
    position = ggplot2::position_dodge(width = 0.45),
    size = 0.9, fatten = 5
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_y_log10(limits = c(y_min_all, y_max_all)) +
  ggplot2::labs(
    title = "Incidence Rate Ratios: Total, Violent, and Non-Violent Crime on Hot/Extreme Days",
    y = "Incidence Rate Ratio (log scale)",
    x = "Heat Index Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "top",
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text = ggplot2::element_text(size = 14, face = "bold"),
    axis.text.x = ggplot2::element_text(angle = 20, hjust = 1)
  )

print(p_all_three)

```



## Create the Final AJPH Figure
```{r}
irr_all_three$city <- dplyr::recode(irr_all_three$city, "Washington, D.C." = "Washington, DC")

irr_all_three$city <- dplyr::recode(irr_all_three$city, "Baltimore" = "Baltimore, MD")

irr_all_three$city <- dplyr::recode(irr_all_three$city, 
  "DC" = "Washington, DC",
  "Baltimore" = "Baltimore, MD"
)

p_all_three <- ggplot2::ggplot(
  irr_all_three,
  ggplot2::aes(
    x = hi_level, y = IRR, ymin = LCL, ymax = UCL,
    shape = outcome, color = outcome
  )
) +
  ggplot2::geom_hline(
    yintercept = 1,
    linetype = "dashed",
    color = "grey40",
    linewidth = 0.6
  ) +
  ggplot2::geom_pointrange(
    position = ggplot2::position_dodge(width = 0.45),
    size = 1.2,      
    fatten = 7       
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_y_log10(limits = c(y_min_all, y_max_all)) +
  ggplot2::scale_x_discrete(
    labels = c(
      "High Heat (85-95th)",
      "Extreme Heat (≥ 95th)"
    )
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Total"       = "#440154",  # dark purple
      "Violent"     = "#31688e",  # blue
      "Non-Violent" = "#35b779"   # green
    ),
    name = "Crime Type"
  ) +
  ggplot2::scale_shape_manual(
    values = c("Total" = 16, "Violent" = 17, "Non-Violent" = 15),
    name = "Crime Type"
  ) +
  ggplot2::labs(
    title = "Incidence Rate Ratios: Total, Violent, and Non-Violent Crime on Hot/Extreme Days",
    y = "Incidence Rate Ratio",
    x = "Heat Index Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
  panel.grid.minor = ggplot2::element_blank(),
  legend.position = "inside",
  legend.position.inside = c(0.91, 0.75),
  legend.background = ggplot2::element_rect(fill = "white", color = "grey50"),
  strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
  strip.text = ggplot2::element_text(size = 12, face = "bold"),
  axis.title.x = ggplot2::element_text(size = 16, color = "black"),
  axis.text.x  = ggplot2::element_text(size = 16, angle = 20, hjust = 1, color = "black"),
  axis.title.y = ggplot2::element_text(size = 14, color = "black"),
  axis.text.y  = ggplot2::element_text(size = 12, color = "black")
)
print(p_all_three)
```

## Save Figures
```{r}
ggplot2::ggsave(
  filename = "AJPH_Fig2_All_Three_Crime_Types.png",
  plot     = p_all_three,
  width    = 10,
  height   = 6,
  units    = "in",
  dpi      = 300)
```

## Create Summary Table
```{r}
# IRR summary table

irr_summary_table <- irr_all_three %>%
  dplyr::mutate(
    # City labels
    city = dplyr::recode(as.character(city), "DC" = "Washington, DC"),

    # Heat category labels
    hi_level = dplyr::recode(
      as.character(hi_level),
      "Hot (85–95th)"   = "Hot (85–95th)",
      "Extreme (≥95th)" = "Extreme Heat (≥95th)"
    ),
    hi_level = factor(hi_level, levels = c("Hot (85–95th)", "Extreme Heat (≥95th)")),

    # Combine IRR + CI into one column
    IRR_CI = sprintf("%.3f (%.3f–%.3f)", IRR, LCL, UCL)
  ) %>%
  dplyr::select(
    City = city,
    `Crime Type` = outcome,
    `Heat Category` = hi_level,
    `IRR (95% CI)` = IRR_CI
  ) %>%
  dplyr::arrange(City, `Crime Type`, `Heat Category`)

# Print table in the knitted HTML
irr_summary_table %>%
  knitr::kable(
    format = "html",
    caption = "Summary of Incidence Rate Ratios by Crime Type and Heat Category"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) %>%
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

# Save CSV
readr::write_csv(irr_summary_table, "AJPH_IRR_Summary_Table.csv")
```

Count of days in each heat category
```{r}
# Count days by heat category
heat_day_counts <- crime_daily %>%
  filter(hi_cat != "Missing") %>%
  group_by(city, hi_cat) %>%
  summarise(
    n_days = n(),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = city, values_from = n_days)

heat_day_counts %>%
  kable(
    format = "html",
    col.names = c("Heat Category", "Baltimore (days)", "DC (days)"),
    caption = "Number of Days in Each Heat Index Category (2020-2024)"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )
```

Summarize HI values for each category
```{r}
# Summary statistics for HI by category
hi_summary_by_cat <- crime_daily %>%
  filter(hi_cat != "Missing") %>%
  group_by(city, hi_cat) %>%
  summarise(
    mean_hi = mean(hi_max_f, na.rm = TRUE),
    median_hi = median(hi_max_f, na.rm = TRUE),
    min_hi = min(hi_max_f, na.rm = TRUE),
    max_hi = max(hi_max_f, na.rm = TRUE),
    sd_hi = sd(hi_max_f, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(across(mean_hi:sd_hi, ~round(., 1)))

hi_summary_by_cat %>%
  kable(
    format = "html",
    col.names = c("City", "Heat Category", "Mean HI (°F)", 
                  "Median HI (°F)", "Min HI (°F)", "Max HI (°F)", "SD"),
    caption = "Summary of Daily Maximum Heat Index by Category"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )
```

## Non-linear Dose-Response Curves (Splines) — AJPH Style
```{r}
# Non-linear dose-response, styled to match p_all_three (Final AJPH Figure)

## Fit spline models (same models as before)
mod_dc_violent_ns    <- MASS::glm.nb(n_violent    ~ splines::ns(hi_max_f, df = 4) + dow + factor(month), data = dc_data)
mod_dc_nonviolent_ns <- MASS::glm.nb(n_nonviolent ~ splines::ns(hi_max_f, df = 4) + dow + factor(month), data = dc_data)

mod_bal_violent_ns    <- MASS::glm.nb(n_violent    ~ splines::ns(hi_max_f, df = 4) + dow + factor(month), data = bal_data)
mod_bal_nonviolent_ns <- MASS::glm.nb(n_nonviolent ~ splines::ns(hi_max_f, df = 4) + dow + factor(month), data = bal_data)

## Prediction grid function (unchanged)
make_pred_grid <- function(data, model, city_label, outcome_label) {

  hi_seq <- seq(
    floor(min(data$hi_max_f, na.rm = TRUE)),
    ceiling(max(data$hi_max_f, na.rm = TRUE)),
    length.out = 200
  )

  newdat <- data.frame(
    hi_max_f = hi_seq,
    dow      = factor("Wed", levels = levels(data$dow)),
    month    = factor("Jul", levels = levels(data$month))
  )

  pred <- predict(model, newdata = newdat, type = "link", se.fit = TRUE)

  newdat %>%
    dplyr::mutate(
      fit      = exp(pred$fit),
      lower    = exp(pred$fit - 1.96 * pred$se.fit),
      upper    = exp(pred$fit + 1.96 * pred$se.fit),
      city     = city_label,
      outcome  = outcome_label
    )
}

pred_dc_violent     <- make_pred_grid(dc_data,  mod_dc_violent_ns,     "DC",        "Violent")
pred_dc_nonviolent  <- make_pred_grid(dc_data,  mod_dc_nonviolent_ns,  "DC",        "Non-Violent")
pred_bal_violent    <- make_pred_grid(bal_data, mod_bal_violent_ns,    "Baltimore", "Violent")
pred_bal_nonviolent <- make_pred_grid(bal_data, mod_bal_nonviolent_ns, "Baltimore", "Non-Violent")

pred_all_ns <- dplyr::bind_rows(
  pred_dc_violent, pred_dc_nonviolent,
  pred_bal_violent, pred_bal_nonviolent
)

## Match city labels to Final AJPH Figure exactly
pred_all_ns$city <- dplyr::recode(pred_all_ns$city,
  "DC"        = "Washington, DC",
  "Baltimore" = "Baltimore, MD"
)

pred_all_ns <- pred_all_ns %>%
  dplyr::mutate(
    city    = factor(city, levels = c("Washington, DC", "Baltimore, MD")),
    outcome = factor(outcome, levels = c("Violent", "Non-Violent"))
  )

## Percentile threshold lines, matched to same city labels
thresh_lines <- hi_thresholds_hw %>%
  dplyr::mutate(
    city = dplyr::recode(city,
      "DC"        = "Washington, DC",
      "Baltimore" = "Baltimore, MD"
    ),
    city = factor(city, levels = c("Washington, DC", "Baltimore, MD"))
  )

## Plot — matched palette, fonts, sizes, legend style from p_all_three
p_dose_response <- ggplot2::ggplot(
  pred_all_ns,
  ggplot2::aes(x = hi_max_f, y = fit, color = outcome, fill = outcome)
) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = lower, ymax = upper), alpha = 0.15, color = NA) +
  ggplot2::geom_line(linewidth = 1.2) +
  ggplot2::geom_vline(
    data = thresh_lines, ggplot2::aes(xintercept = p85),
    linetype = "dashed", color = "grey40", linewidth = 0.6
  ) +
  ggplot2::geom_vline(
    data = thresh_lines, ggplot2::aes(xintercept = p95),
    linetype = "dotted", color = "grey20", linewidth = 0.6
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_color_manual(
    values = c(
      "Violent"     = "#31688e",
      "Non-Violent" = "#35b779"
    ),
    name = "Crime Type"
  ) +
  ggplot2::scale_fill_manual(
    values = c(
      "Violent"     = "#31688e",
      "Non-Violent" = "#35b779"
    ),
    name = "Crime Type"
  ) +
  ggplot2::labs(
    title = "Non-Linear Association Between Daily Maximum Heat Index and Crime",
    y = "Predicted Daily Crime Count",
    x = "Daily Maximum Heat Index (°F)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.96, 0.65),
    legend.background = ggplot2::element_rect(fill = "white", color = "grey50"),
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text = ggplot2::element_text(size = 12, face = "bold"),
    axis.title.x = ggplot2::element_text(size = 16, color = "black"),
    axis.text.x  = ggplot2::element_text(size = 16, color = "black"),
    axis.title.y = ggplot2::element_text(size = 14, color = "black"),
    axis.text.y  = ggplot2::element_text(size = 12, color = "black")
  )

print(p_dose_response)

ggplot2::ggsave(
  filename = "AJPH_DoseResponse_Spline_Crime_vs_HeatIndex.png",
  plot     = p_dose_response,
  width    = 10,
  height   = 6,
  units    = "in",
  dpi      = 300
)
```

```{r}
# Find the HI value at the peak of the fitted curve (Baltimore non-violent)
peak_row <- pred_bal_nonviolent %>%
  dplyr::filter(fit == max(fit))

peak_hi <- peak_row$hi_max_f

# Choose a low reference point - e.g., the 10th percentile of observed HI,
# or the minimum of your prediction grid
low_hi <- min(pred_bal_nonviolent$hi_max_f)

# Build a small newdata frame with just these two points
newdat_contrast <- data.frame(
  hi_max_f = c(low_hi, peak_hi),
  dow      = factor("Wed", levels = levels(bal_data$dow)),
  month    = factor("Jul", levels = levels(bal_data$month))
)

# Get predictions on the link scale with SEs
pred_contrast <- predict(mod_bal_nonviolent_ns, newdata = newdat_contrast,
                          type = "link", se.fit = TRUE)


library(emmeans)

# emmeans approach: get estimated marginal means at the two HI values, then contrast them
emm <- emmeans::emmeans(mod_bal_nonviolent_ns, ~ hi_max_f,
                         at = list(hi_max_f = c(low_hi, peak_hi)),
                         type = "response")
print(emm)

# Now test the contrast (ratio) between peak and low point
contrast_result <- emmeans::contrast(emm, method = "revpairwise", type = "response")
print(contrast_result)
confint(contrast_result)
```


# Sensitivity Analysis

## Workday vs. Day off Work (DOW) analysis
```{r}
# Create workday variable
crime_daily <- crime_daily %>%
  mutate(
    workday = ifelse(dow %in% c("Sat", "Sun"), "Weekend", "Workday"),
    workday = factor(workday, levels = c("Workday", "Weekend"))
  )

# Recreate the city-specific datasets AFTER adding workday
crime_daily_dc <- crime_daily %>%
  dplyr::filter(city == "DC")

crime_daily_bal <- crime_daily %>%
  dplyr::filter(city == "Baltimore")

# NOW refit models with workday instead of dow
dc_data_wd  <- crime_daily_dc  %>% filter(hi_cat != "Missing")
bal_data_wd <- crime_daily_bal %>% filter(hi_cat != "Missing")

# DC - workday models
mod_dc_total_wd      <- MASS::glm.nb(n_total      ~ hi_cat + workday + factor(month), data = dc_data_wd)
mod_dc_violent_wd    <- MASS::glm.nb(n_violent    ~ hi_cat + workday + factor(month), data = dc_data_wd)
mod_dc_nonviolent_wd <- MASS::glm.nb(n_nonviolent ~ hi_cat + workday + factor(month), data = dc_data_wd)

# Baltimore - workday models
mod_bal_total_wd      <- MASS::glm.nb(n_total      ~ hi_cat + workday + factor(month), data = bal_data_wd)
mod_bal_violent_wd    <- MASS::glm.nb(n_violent    ~ hi_cat + workday + factor(month), data = bal_data_wd)
mod_bal_nonviolent_wd <- MASS::glm.nb(n_nonviolent ~ hi_cat + workday + factor(month), data = bal_data_wd)

# Compare IRRs for heat categories between models
# Extract IRRs from workday models
irr_dc_wd <- dplyr::bind_rows(
  irr_df(mod_dc_total_wd,      "Total",       "DC"),
  irr_df(mod_dc_violent_wd,    "Violent",     "DC"),
  irr_df(mod_dc_nonviolent_wd, "Non-Violent", "DC")
)

irr_bal_wd <- dplyr::bind_rows(
  irr_df(mod_bal_total_wd,      "Total",       "Baltimore"),
  irr_df(mod_bal_violent_wd,    "Violent",     "Baltimore"),
  irr_df(mod_bal_nonviolent_wd, "Non-Violent", "Baltimore")
)

# Compare side-by-side
comparison <- irr_dc %>%
  dplyr::select(city, outcome, hi_level, irr_dow = irr) %>%
  dplyr::left_join(
    irr_dc_wd %>% dplyr::select(city, outcome, hi_level, irr_workday = irr),
    by = c("city", "outcome", "hi_level")
  )


print(comparison)
```

### Include holidays in DOW:

##### I got the DC holiday dates from: https://edpm.dc.gov/issuances/legal-public-holidays-2023/ (for every different year)

DC Holidays in 2020:
- New Year’s Day	Wednesday, January 1
- Martin Luther King Jr’s Birthday	Monday, January 20
- Washington’s Birthday	Monday, February 17
- D.C. Emancipation Day	Thursday, April 16
- Memorial Day	Monday, May 25
- Independence Day	Friday, July 3
- Labor Day	Monday, September 7
- Indigenous Peoples’ Day	Monday, October 12
- Veterans Day	Wednesday, November 11
- Thanksgiving Day	Thursday, November 26
- Christmas Day	Friday, December 25

DC Holidays in 2021:
- New Year's Day	Friday, January 1, 2021
- Dr. Martin Luther King Jr.'s Birthday	Monday, January 18, 2021
- Inauguration Day	Wednesday, January 20, 2021
- Washington's Birthday	Monday, February 15, 2021
- D.C. Emancipation Day	Friday, April 16, 2021
- Memorial Day	Monday, May 31, 2021
- Juneteenth National Independence Day	Friday, June 18, 2021
- Independence Day	Monday, July 5, 2021
- Labor Day	Monday, September 6, 2021
- Indigenous Peoples' Day	Monday, October 11, 2021
- Veterans Day	Thursday, November 11, 2021
- Thanksgiving Day	Thursday, November 25, 2021
- Christmas Day	Friday, December 24, 2021

DC Holidays in 2022:
- New Year's Day	Friday, December 31, 2021
- Dr. Martin Luther King, Jr.'s Birthday	Monday, January 17, 2022
- Washington's Birthday	Monday, February 21, 2022
- D.C. Emancipation Day	Friday, April 15, 2022
- Memorial Day	Monday, May 30, 2022
- Juneteenth National Independence Day	Monday, June 20, 2022
- Independence Day	Monday, July 4, 2022
- Labor Day	Monday, September 5, 2022
- Indigenous Peoples' Day	Monday, October 10, 2022
- Veterans Day	Friday, November 11, 2022
- Thanksgiving Day	Thursday, November 24, 2022
- Christmas Day	Monday, December 26, 2022

DC Holidays in 2023:
- New Year's Day	Monday, January 2, 2023
- Dr. Martin Luther King, Jr.'s Birthday	Monday, January 16, 2023
- Washington's Birthday	Monday, February 20, 2023
- D.C. Emancipation Day	Monday, April 17, 2023
- Memorial Day	Monday, May 29, 2023
- Juneteenth National Independence Day	Monday, June 19, 2023
- Independence Day	Tuesday, July 4, 2023
- Labor Day	Monday, September 4, 2023
- Indigenous Peoples' Day	Monday, October 9, 2023
- Veterans Day	Friday, November 10, 2023
- Thanksgiving Day	Thursday, November 23, 2023
- Christmas Day	Monday, December 25, 2023


DC Holidays in 2024:
- New Year's Day	Monday, January 1, 2024
- Dr. Martin Luther King, Jr.'s Birthday	Monday, January 15, 2024
- Washington's Birthday	Monday, February 19, 2024
- D.C. Emancipation Day	Tuesday, April 16, 2024
- Memorial Day	Monday, May 27, 2024
- Juneteenth National Independence Day	Wednesday, June 19, 2024
- Independence Day	Thursday, July 4, 2024
- Labor Day	Monday, September 2, 2024
- Indigenous Peoples' Day	Monday, October 14, 2024
- Veterans Day	Monday, November 11, 2024
- Thanksgiving Day	Thursday, November 28, 2024
- Christmas Day	Wednesday, December 25, 2024

#### Maryland public holidays: 
I retrieved Maryland public holidays from: https://dbm.maryland.gov/employees/Pages/StateHolidays2023.aspx

Maryland Holidays in 2020:
New Year's Day	Wednesday	January 1
Birthday of Dr. Martin Luther King, Jr	Monday	January 20
Presidents' Day	Monday	February 17
Memorial Day	Monday	May 25
Independence Day	Friday	July 3
Labor Day	Monday	September 7
Columbus Day	Monday	October 12
Election Day	Tuesday	November 3
Veterans Day	Wednesday November 11
Thanksgiving Day	Thursday	November 26
American Indian Heritage Day	Friday	November 27
Christmas Day	Friday	December 25

Maryland Holidays in 2021:
- New Year's Day Friday	January 1 
- Birthday of Dr. Martin Luther King, Jr	Monday	January 18
- Presidents' Day	Monday	February 15
- Memorial Day	Monday	May 31 Independence Day	Monday	July 5
- Labor Day	Monday	September 6
- Columbus Day	Monday	October 11
- Veterans Day	Thursday November 11
- Thanksgiving Day	Thursday	November 25
- American Indian Heritage Day	Friday	November 26
- Christmas Day	Friday	December 24
- New Year's Day	Friday	December 31

Maryland Holidays in 2022:
- New Year's Day*	Friday	December 31 
- Birthday of Dr. Martin Luther King, Jr	Monday	January 17 
- Presidents' Day	Monday	February 21 
- Memorial Day	Monday	May 30 Juneteenth	Monday	June 20 
- Independence Day	Monday	July 4 
- Labor Day	Monday	September 5
- Columbus Day	Monday	October 10
- Gubernatorial Election Day	Tuesday	November 8
- Veterans Day	Friday November 11
- Thanksgiving Day	Thursday	November 24
- American Indian Heritage Day	Friday	November 25
- Christmas Day*	Monday	December 26

Maryland Holidays in 2023:
- New Year's Day Monday January 2 
- Birthday of Dr. Martin Luther King, Jr. Monday January 16 
- Presidents' Day	Monday	February 20 
- Memorial Day	Monday	May 29 
- Juneteenth Monday June 19 
- Independence Day	Tuesday July 4 
- Labor Day	Monday	September 4
- Columbus Day	Monday	October 9 
- Veterans' Day Friday November 10 
- Thanksgiving Day	Thursday	November 23 
- American Indian Heritage Day	Friday	November 24 
- Christmas Day Monday December 25

Maryland Holidays in 2024:
- New Year's Day Monday January 1
- Birthday of Dr. Martin Luther King, Jr. Monday January 15
- Presidents' Day	Monday	February 19 
- Memorial Day	Monday	May 27 
- Juneteenth Wednesday June 19 
- Independence Day	Thursday July 4 
- Labor Day	Monday	September 2 
- Columbus Day	Monday	October 14 
- Election Day Tuesday November 5 
- Veterans' Day Monday November 11 
- Thanksgiving Day	Thursday	November 28 
- American Indian Heritage Day	Friday	November 29 
- Christmas Day Wednesday December 25

### Another sensivity Analysis Workday vs DOW (Weekend + Holiday)
```{r}
# Sensitivity Analysis: Workday vs. Non-Workday (including holidays)

## Define DC holidays
dc_holidays <- as.Date(c(
  # 2020
  "2020-01-01", "2020-01-20", "2020-02-17", "2020-04-16", "2020-05-25",
  "2020-07-03", "2020-09-07", "2020-10-12", "2020-11-11", "2020-11-26",
  "2020-12-25",
  # 2021
  "2021-01-01", "2021-01-18", "2021-01-20", "2021-02-15", "2021-04-16",
  "2021-05-31", "2021-06-18", "2021-07-05", "2021-09-06", "2021-10-11",
  "2021-11-11", "2021-11-25", "2021-12-24",
  # 2022
  "2021-12-31", "2022-01-17", "2022-02-21", "2022-04-15", "2022-05-30",
  "2022-06-20", "2022-07-04", "2022-09-05", "2022-10-10", "2022-11-11",
  "2022-11-24", "2022-12-26",
  # 2023
  "2023-01-02", "2023-01-16", "2023-02-20", "2023-04-17", "2023-05-29",
  "2023-06-19", "2023-07-04", "2023-09-04", "2023-10-09", "2023-11-10",
  "2023-11-23", "2023-12-25",
  # 2024
  "2024-01-01", "2024-01-15", "2024-02-19", "2024-04-16", "2024-05-27",
  "2024-06-19", "2024-07-04", "2024-09-02", "2024-10-14", "2024-11-11",
  "2024-11-28", "2024-12-25"
))

## Define Maryland holidays
md_holidays <- as.Date(c(
  # 2020
  "2020-01-01", "2020-01-20", "2020-02-17", "2020-05-25", "2020-07-03",
  "2020-09-07", "2020-10-12", "2020-11-03", "2020-11-11", "2020-11-26",
  "2020-11-27", "2020-12-25",
  # 2021
  "2021-01-01", "2021-01-18", "2021-02-15", "2021-05-31", "2021-07-05",
  "2021-09-06", "2021-10-11", "2021-11-11", "2021-11-25", "2021-11-26",
  "2021-12-24", "2021-12-31",
  # 2022
  "2022-01-17", "2022-02-21", "2022-05-30", "2022-06-20",
  "2022-07-04", "2022-09-05", "2022-10-10", "2022-11-08", "2022-11-11",
  "2022-11-24", "2022-11-25", "2022-12-26",
  # 2023
  "2023-01-02", "2023-01-16", "2023-02-20", "2023-05-29", "2023-06-19",
  "2023-07-04", "2023-09-04", "2023-10-09", "2023-11-10", "2023-11-23",
  "2023-11-24", "2023-12-25",
  # 2024
  "2024-01-01", "2024-01-15", "2024-02-19", "2024-05-27", "2024-06-19",
  "2024-07-04", "2024-09-02", "2024-10-14", "2024-11-05", "2024-11-11",
  "2024-11-28", "2024-11-29", "2024-12-25"
))

## Create workday variable (weekends + holidays = non-workday)
crime_daily <- crime_daily %>%
  mutate(
    # First check if weekend
    is_weekend = dow %in% c("Sat", "Sun"),
    # Then check if holiday (city-specific)
    is_holiday = case_when(
      city == "DC" ~ crime_date %in% dc_holidays,
      city == "Baltimore" ~ crime_date %in% md_holidays,
      TRUE ~ FALSE
    ),
    # Combine weekend and holiday
    workday_incl_holidays = case_when(
      is_weekend | is_holiday ~ "Non-Workday",
      TRUE ~ "Workday"
    ),
    workday_incl_holidays = factor(workday_incl_holidays, 
                                   levels = c("Workday", "Non-Workday"))
  )

# Verify the number of holidays captured
cat("DC holidays in dataset:", 
    sum(crime_daily$city == "DC" & crime_daily$is_holiday, na.rm = TRUE), "\n")
cat("Baltimore holidays in dataset:", 
    sum(crime_daily$city == "Baltimore" & crime_daily$is_holiday, na.rm = TRUE), "\n")

# Summary of workdays vs non-workdays by city
crime_daily %>%
  group_by(city, workday_incl_holidays) %>%
  summarise(n_days = n(), .groups = "drop") %>%
  kable(
    format = "html",
    col.names = c("City", "Day Type", "Number of Days"),
    caption = "Distribution of Workdays vs Non-Workdays (including holidays)"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )

# RECREATE the city-specific datasets AFTER adding workday_incl_holidays
crime_daily_dc <- crime_daily %>%
  dplyr::filter(city == "DC")

crime_daily_bal <- crime_daily %>%
  dplyr::filter(city == "Baltimore")

# Refit models with workday_incl_holidays
dc_data_wdh  <- crime_daily_dc  %>% filter(hi_cat != "Missing")
bal_data_wdh <- crime_daily_bal %>% filter(hi_cat != "Missing")

# DC - workday models (including holidays)
mod_dc_total_wdh      <- MASS::glm.nb(n_total      ~ hi_cat + workday_incl_holidays + factor(month), 
                                      data = dc_data_wdh)
mod_dc_violent_wdh    <- MASS::glm.nb(n_violent    ~ hi_cat + workday_incl_holidays + factor(month), 
                                      data = dc_data_wdh)
mod_dc_nonviolent_wdh <- MASS::glm.nb(n_nonviolent ~ hi_cat + workday_incl_holidays + factor(month), 
                                      data = dc_data_wdh)

# Baltimore - workday models (including holidays)
mod_bal_total_wdh      <- MASS::glm.nb(n_total      ~ hi_cat + workday_incl_holidays + factor(month), 
                                       data = bal_data_wdh)
mod_bal_violent_wdh    <- MASS::glm.nb(n_violent    ~ hi_cat + workday_incl_holidays + factor(month), 
                                       data = bal_data_wdh)
mod_bal_nonviolent_wdh <- MASS::glm.nb(n_nonviolent ~ hi_cat + workday_incl_holidays + factor(month), 
                                       data = bal_data_wdh)

# Extract IRRs from workday+holidays models
irr_dc_wdh <- dplyr::bind_rows(
  irr_df(mod_dc_total_wdh,      "Total",       "DC"),
  irr_df(mod_dc_violent_wdh,    "Violent",     "DC"),
  irr_df(mod_dc_nonviolent_wdh, "Non-Violent", "DC")
)

irr_bal_wdh <- dplyr::bind_rows(
  irr_df(mod_bal_total_wdh,      "Total",       "Baltimore"),
  irr_df(mod_bal_violent_wdh,    "Violent",     "Baltimore"),
  irr_df(mod_bal_nonviolent_wdh, "Non-Violent", "Baltimore")
)

# Compare all three approaches: DOW vs Workday vs Workday+Holidays
comparison_all <- irr_dc %>%
  dplyr::select(city, outcome, hi_level, irr_dow = irr, 
                lower_dow = lower.CL, upper_dow = upper.CL) %>%
  dplyr::left_join(
    irr_dc_wd %>% dplyr::select(city, outcome, hi_level, irr_workday = irr,
                                lower_wd = lower.CL, upper_wd = upper.CL),
    by = c("city", "outcome", "hi_level")
  ) %>%
  dplyr::left_join(
    irr_dc_wdh %>% dplyr::select(city, outcome, hi_level, irr_workday_holidays = irr,
                                 lower_wdh = lower.CL, upper_wdh = upper.CL),
    by = c("city", "outcome", "hi_level")
  )

# Add Baltimore data
comparison_all <- comparison_all %>%
  dplyr::bind_rows(
    irr_bal %>%
      dplyr::select(city, outcome, hi_level, irr_dow = irr, 
                    lower_dow = lower.CL, upper_dow = upper.CL) %>%
      dplyr::left_join(
        irr_bal_wd %>% dplyr::select(city, outcome, hi_level, irr_workday = irr,
                                     lower_wd = lower.CL, upper_wd = upper.CL),
        by = c("city", "outcome", "hi_level")
      ) %>%
      dplyr::left_join(
        irr_bal_wdh %>% dplyr::select(city, outcome, hi_level, irr_workday_holidays = irr,
                                      lower_wdh = lower.CL, upper_wdh = upper.CL),
        by = c("city", "outcome", "hi_level")
      )
  )

# Create formatted comparison table
comparison_table <- comparison_all %>%
  mutate(
    DOW_Model = sprintf("%.2f (%.2f–%.2f)", irr_dow, lower_dow, upper_dow),
    Workday_Model = sprintf("%.2f (%.2f–%.2f)", irr_workday, lower_wd, upper_wd),
    Workday_Holidays_Model = sprintf("%.2f (%.2f–%.2f)", irr_workday_holidays, lower_wdh, upper_wdh)
  ) %>%
  dplyr::select(city, outcome, hi_level, DOW_Model, Workday_Model, Workday_Holidays_Model)

# Display comparison table
comparison_table %>%
  kable(
    format = "html",
    col.names = c("City", "Crime Type", "Heat Level", 
                  "DOW Model IRR (95% CI)", 
                  "Workday Model IRR (95% CI)",
                  "Workday + Holidays Model IRR (95% CI)"),
    caption = "Sensitivity Analysis: Comparison of IRRs Across Different Day-of-Week Specifications"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) %>%
  collapse_rows(columns = 1:2, valign = "top")

# Save comparison table
readr::write_csv(comparison_table, "sensitivity_analysis_dow_comparison.csv")

# Print summary
cat("\n=== SENSITIVITY ANALYSIS SUMMARY ===\n")
cat("Models fitted:\n")
cat("1. DOW Model: Adjusts for day of week (7 categories)\n")
cat("2. Workday Model: Adjusts for workday vs weekend (2 categories)\n")
cat("3. Workday + Holidays Model: Adjusts for workday vs non-workday (weekends + holidays)\n\n")
```

### Another Sensitivity Analysis: Max Temp
Using daily maximum temperature instead of heat index
```{r}
# Sensitivity Analysis: Daily Maximum Temperature Instead of Heat Index

## Extract Daily Maximum Temperature from ASOS data
dc_daily_temp <- dc_hr %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(
    temp_mean_f = if (all(is.na(tmpf))) NA_real_ else mean(tmpf, na.rm = TRUE),
    temp_max_f  = if (all(is.na(tmpf))) NA_real_ else max(tmpf,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(date <= as.Date("2024-12-31")) %>%
  dplyr::mutate(city = "DC")

bal_daily_temp <- bal_hr %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(
    temp_mean_f = if (all(is.na(tmpf))) NA_real_ else mean(tmpf, na.rm = TRUE),
    temp_max_f  = if (all(is.na(tmpf))) NA_real_ else max(tmpf,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(date <= as.Date("2024-12-31")) %>%
  dplyr::mutate(city = "Baltimore")

# Combine temperature data
daily_temp_all <- dplyr::bind_rows(dc_daily_temp, bal_daily_temp) %>%
  dplyr::select(city, date, temp_mean_f, temp_max_f)

## Calculate Temperature Percentile Thresholds (85th and 95th)
temp_thresholds <- daily_temp_all %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    p85_temp = quantile(temp_max_f, 0.85, na.rm = TRUE),
    p95_temp = quantile(temp_max_f, 0.95, na.rm = TRUE),
    .groups = "drop"
  )

# Display thresholds
temp_thresholds %>%
  dplyr::mutate(
    p85_temp = round(p85_temp, 1),
    p95_temp = round(p95_temp, 1)
  ) %>%
  knitr::kable(
    format = "html",
    col.names = c("City", "85th Percentile (Max Temp °F)", "95th Percentile (Max Temp °F)"),
    caption = "Temperature Thresholds (Daily Max Temperature)"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )

## Create Crime Daily Dataset with Temperature Categories
crime_daily_temp <- crime_all %>%
  dplyr::mutate(
    dow   = lubridate::wday(crime_date, label = TRUE, abbr = TRUE),
    month = lubridate::month(crime_date, label = TRUE, abbr = TRUE),
    year  = lubridate::year(crime_date)
  ) %>%
  dplyr::group_by(city, crime_date, dow, month, year) %>%
  dplyr::summarise(
    n_total      = dplyr::n(),
    n_violent    = sum(toupper(as.character(crime_type)) == "VIOLENT", na.rm = TRUE),
    n_nonviolent = n_total - n_violent,
    .groups = "drop"
  ) %>%
  dplyr::left_join(daily_temp_all, by = c("city", "crime_date" = "date")) %>%
  dplyr::left_join(temp_thresholds, by = "city") %>%
  dplyr::mutate(
    temp_cat = dplyr::case_when(
      is.na(temp_max_f)                        ~ "Missing",
      temp_max_f <  p85_temp                   ~ "<85th",
      temp_max_f >= p85_temp & temp_max_f < p95_temp ~ "Hot (85–95th)",
      temp_max_f >= p95_temp                   ~ "Extreme (≥95th)"
    ),
    temp_cat = factor(temp_cat, levels = c("<85th","Hot (85–95th)","Extreme (≥95th)","Missing"))
  )

# Split by city
crime_daily_temp_dc <- crime_daily_temp %>%
  dplyr::filter(city == "DC")

crime_daily_temp_bal <- crime_daily_temp %>%
  dplyr::filter(city == "Baltimore")

## Fit Negative Binomial Models Using Temperature
dc_data_temp  <- crime_daily_temp_dc  %>% dplyr::filter(temp_cat != "Missing")
bal_data_temp <- crime_daily_temp_bal %>% dplyr::filter(temp_cat != "Missing")

# DC models
mod_dc_total_temp      <- MASS::glm.nb(n_total      ~ temp_cat + dow + factor(month), data = dc_data_temp)
mod_dc_violent_temp    <- MASS::glm.nb(n_violent    ~ temp_cat + dow + factor(month), data = dc_data_temp)
mod_dc_nonviolent_temp <- MASS::glm.nb(n_nonviolent ~ temp_cat + dow + factor(month), data = dc_data_temp)

# Baltimore models
mod_bal_total_temp      <- MASS::glm.nb(n_total      ~ temp_cat + dow + factor(month), data = bal_data_temp)
mod_bal_violent_temp    <- MASS::glm.nb(n_violent    ~ temp_cat + dow + factor(month), data = bal_data_temp)
mod_bal_nonviolent_temp <- MASS::glm.nb(n_nonviolent ~ temp_cat + dow + factor(month), data = bal_data_temp)

## Modified IRR helper function that accepts variable name
irr_df_custom <- function(model, outcome_label, city_label, var_name = "hi_cat") {
  
  # Create formula dynamically
  em <- emmeans::emmeans(model, stats::as.formula(paste("~", var_name)), type = "response")
  ct <- emmeans::contrast(em, method = "trt.vs.ctrl", ref = 1, type = "response")
  
  ct_df <- as.data.frame(stats::confint(ct))
  
  # Harmonize CI column names
  nm <- names(ct_df)
  if (!"lower.CL" %in% nm) {
    if ("asymp.LCL" %in% nm) names(ct_df)[nm == "asymp.LCL"] <- "lower.CL"
    else if ("LCL" %in% nm)  names(ct_df)[nm == "LCL"]       <- "lower.CL"
  }
  nm <- names(ct_df)
  if (!"upper.CL" %in% nm) {
    if ("asymp.UCL" %in% nm) names(ct_df)[nm == "asymp.UCL"] <- "upper.CL"
    else if ("UCL" %in% nm)  names(ct_df)[nm == "UCL"]       <- "upper.CL"
  }
  
  # IRR column (usually "ratio")
  ycol <- if ("ratio" %in% names(ct_df)) "ratio" else {
    cand <- intersect(c("response", "estimate", "emmean"), names(ct_df))
    if (length(cand) == 0) stop("Couldn't find IRR/estimate column in ct_df.")
    cand[1]
  }
  
  ct_df %>%
    dplyr::mutate(
      hi_level = sub(paste0(" / <85th"), "", contrast),
      irr      = .data[[ycol]],
      outcome  = outcome_label,
      city     = city_label
    ) %>%
    dplyr::select(city, outcome, hi_level, irr, lower.CL, upper.CL)
}

## Extract IRRs for Temperature Models
irr_dc_temp <- dplyr::bind_rows(
  irr_df_custom(mod_dc_total_temp,      "Total",       "DC", "temp_cat"),
  irr_df_custom(mod_dc_violent_temp,    "Violent",     "DC", "temp_cat"),
  irr_df_custom(mod_dc_nonviolent_temp, "Non-Violent", "DC", "temp_cat")
)

irr_bal_temp <- dplyr::bind_rows(
  irr_df_custom(mod_bal_total_temp,      "Total",       "Baltimore", "temp_cat"),
  irr_df_custom(mod_bal_violent_temp,    "Violent",     "Baltimore", "temp_cat"),
  irr_df_custom(mod_bal_nonviolent_temp, "Non-Violent", "Baltimore", "temp_cat")
)

# Combine all temperature IRRs
# Combine all temperature IRRs (define it first!)
irr_all_temp <- dplyr::bind_rows(irr_dc_temp, irr_bal_temp) %>%
  dplyr::mutate(
    # Standardize join keys NOW (same as irr_all_three)
    city = dplyr::recode(city,
      "Baltimore" = "Baltimore, MD",
      "DC"        = "DC"
    ),
    hi_level = stringr::str_squish(hi_level),
    hi_level = dplyr::case_when(
      stringr::str_detect(hi_level, "Hot")     ~ "Hot (85–95th)",
      stringr::str_detect(hi_level, "Extreme") ~ "Extreme (≥95th)",
      TRUE ~ hi_level
    )
  ) %>%
  dplyr::rename(IRR = irr, LCL = lower.CL, UCL = upper.CL)


## Compare Heat Index vs Temperature Models
# Combine HI and Temp results for comparison
comparison_hi_vs_temp <- irr_all_three %>%
  dplyr::select(city, outcome, hi_level, IRR_HI = IRR, LCL_HI = LCL, UCL_HI = UCL) %>%
  dplyr::left_join(
    irr_all_temp %>% dplyr::select(city, outcome, hi_level, IRR_Temp = IRR, LCL_Temp = LCL, UCL_Temp = UCL),
    by = c("city", "outcome", "hi_level")
  ) %>%
  dplyr::mutate(
    HI_Model = sprintf("%.3f (%.3f–%.3f)", IRR_HI, LCL_HI, UCL_HI),
    Temp_Model = sprintf("%.3f (%.3f–%.3f)", IRR_Temp, LCL_Temp, UCL_Temp)
  )

# Display comparison table
comparison_hi_vs_temp %>%
  dplyr::select(City = city, `Crime Type` = outcome, `Heat Category` = hi_level, 
                `Heat Index Model IRR (95% CI)` = HI_Model, 
                `Temperature Model IRR (95% CI)` = Temp_Model) %>%
  kable(
    format = "html",
    caption = "Sensitivity Analysis: Comparison of Heat Index vs Maximum Temperature Models"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) %>%
  collapse_rows(columns = 1:2, valign = "top")

# Save comparison
readr::write_csv(comparison_hi_vs_temp, "sensitivity_analysis_HI_vs_Temp.csv")

## Visualize Temperature Model Results
y_min_temp <- min(irr_all_temp$LCL, na.rm = TRUE) * 0.95
y_max_temp <- max(irr_all_temp$UCL, na.rm = TRUE) * 1.05

p_temp_model <- ggplot2::ggplot(
  irr_all_temp,
  ggplot2::aes(
    x = hi_level, y = IRR, ymin = LCL, ymax = UCL,
    shape = outcome, linetype = outcome, color = outcome
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, color = "grey50") +
  ggplot2::geom_pointrange(
    position = ggplot2::position_dodge(width = 0.45),
    size = 1.2,      
    fatten = 7       
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_y_log10(limits = c(y_min_temp, y_max_temp)) +
  ggplot2::scale_color_manual(
    values = c(
      "Total"       = "#440154",
      "Violent"     = "#31688e",
      "Non-Violent" = "#35b779"
    ),
    name = "Crime Type"
  ) +
  ggplot2::scale_shape_manual(
    values = c("Total" = 16, "Violent" = 17, "Non-Violent" = 15),
    name = "Crime Type"
  ) +
  ggplot2::scale_linetype_discrete(name = "Crime Type") +
  ggplot2::labs(
    title = "IRRs Using Daily Maximum Temperature (Sensitivity Analysis)",
    y = "Incidence Rate Ratio",
    x = "Temperature Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "top",
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text = ggplot2::element_text(size = 12, face = "bold"),
    axis.title.x = ggplot2::element_text(size = 14),
    axis.text.x  = ggplot2::element_text(size = 13, angle = 20, hjust = 1),
    axis.title.y = ggplot2::element_text(size = 14),
    axis.text.y  = ggplot2::element_text(size = 12)
  )

print(p_temp_model)

# Save temperature model figure
ggplot2::ggsave(
  filename = "Sensitivity_Temperature_Model.png",
  plot     = p_temp_model,
  width    = 10,
  height   = 6,
  units    = "in",
  dpi      = 300
)

## Summary Statistics
cat("\n=== SENSITIVITY ANALYSIS: TEMPERATURE VS HEAT INDEX ===\n\n")
cat("Temperature Thresholds:\n")
print(temp_thresholds)
cat("\nMissing Temperature values:", sum(is.na(crime_daily_temp$temp_max_f)), "\n")
cat("Date range:", min(crime_daily_temp$crime_date, na.rm = TRUE), "to",
    max(crime_daily_temp$crime_date, na.rm = TRUE), "\n\n")
```

### Another Sensitivity Analysis: Year Effects
Adding year as a covariate
```{r}
# Sensitivity Analysis: Adding Year as a Covariate

## Refit primary models with factor(year) added
dc_data_yr  <- crime_daily_dc  %>% dplyr::filter(hi_cat != "Missing")
bal_data_yr <- crime_daily_bal %>% dplyr::filter(hi_cat != "Missing")

# DC - year-adjusted models
mod_dc_total_yr      <- MASS::glm.nb(n_total      ~ hi_cat + dow + factor(month) + factor(year), data = dc_data_yr)
mod_dc_violent_yr    <- MASS::glm.nb(n_violent    ~ hi_cat + dow + factor(month) + factor(year), data = dc_data_yr)
mod_dc_nonviolent_yr <- MASS::glm.nb(n_nonviolent ~ hi_cat + dow + factor(month) + factor(year), data = dc_data_yr)

# Baltimore - year-adjusted models
mod_bal_total_yr      <- MASS::glm.nb(n_total      ~ hi_cat + dow + factor(month) + factor(year), data = bal_data_yr)
mod_bal_violent_yr    <- MASS::glm.nb(n_violent    ~ hi_cat + dow + factor(month) + factor(year), data = bal_data_yr)
mod_bal_nonviolent_yr <- MASS::glm.nb(n_nonviolent ~ hi_cat + dow + factor(month) + factor(year), data = bal_data_yr)

## Extract IRRs
irr_dc_yr <- dplyr::bind_rows(
  irr_df(mod_dc_total_yr,      "Total",       "DC"),
  irr_df(mod_dc_violent_yr,    "Violent",     "DC"),
  irr_df(mod_dc_nonviolent_yr, "Non-Violent", "DC")
)

irr_bal_yr <- dplyr::bind_rows(
  irr_df(mod_bal_total_yr,      "Total",       "Baltimore"),
  irr_df(mod_bal_violent_yr,    "Violent",     "Baltimore"),
  irr_df(mod_bal_nonviolent_yr, "Non-Violent", "Baltimore")
)

irr_all_yr <- dplyr::bind_rows(irr_dc_yr, irr_bal_yr) %>%
  dplyr::mutate(
    city = dplyr::recode(city,
      "Baltimore" = "Baltimore, MD",
      "DC"        = "DC"
    ),
    hi_level = stringr::str_squish(hi_level),
    hi_level = dplyr::case_when(
      stringr::str_detect(hi_level, "Hot")     ~ "Hot (85–95th)",
      stringr::str_detect(hi_level, "Extreme") ~ "Extreme (≥95th)",
      TRUE ~ hi_level
    )
  ) %>%
  dplyr::rename(IRR = irr, LCL = lower.CL, UCL = upper.CL)

## Compare primary models vs year-adjusted models
comparison_yr <- irr_all_three %>%
  dplyr::select(city, outcome, hi_level, IRR_primary = IRR, LCL_primary = LCL, UCL_primary = UCL) %>%
  dplyr::left_join(
    irr_all_yr %>% dplyr::select(city, outcome, hi_level, IRR_yr = IRR, LCL_yr = LCL, UCL_yr = UCL),
    by = c("city", "outcome", "hi_level")
  ) %>%
  dplyr::mutate(
    Primary_Model    = sprintf("%.3f (%.3f–%.3f)", IRR_primary, LCL_primary, UCL_primary),
    Year_Adj_Model   = sprintf("%.3f (%.3f–%.3f)", IRR_yr,      LCL_yr,      UCL_yr)
  )

## Display comparison table
comparison_yr %>%
  dplyr::select(
    City           = city,
    `Crime Type`   = outcome,
    `Heat Category` = hi_level,
    `Primary Model IRR (95% CI)`       = Primary_Model,
    `Year-Adjusted Model IRR (95% CI)` = Year_Adj_Model
  ) %>%
  knitr::kable(
    format = "html",
    caption = "Sensitivity Analysis: Primary Models vs Year-Adjusted Models"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) %>%
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

## Save comparison
readr::write_csv(comparison_yr, "sensitivity_analysis_year_adjusted.csv")

## Visualize year-adjusted model results
y_min_yr <- min(irr_all_yr$LCL, na.rm = TRUE) * 0.95
y_max_yr <- max(irr_all_yr$UCL, na.rm = TRUE) * 1.05

p_yr_model <- ggplot2::ggplot(
  irr_all_yr,
  ggplot2::aes(
    x = hi_level, y = IRR, ymin = LCL, ymax = UCL,
    shape = outcome, color = outcome
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, color = "grey50") +
  ggplot2::geom_pointrange(
    position = ggplot2::position_dodge(width = 0.45),
    size = 1.2,
    fatten = 7
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_y_log10(limits = c(y_min_yr, y_max_yr)) +
  ggplot2::scale_color_manual(
    values = c(
      "Total"       = "#440154",
      "Violent"     = "#31688e",
      "Non-Violent" = "#35b779"
    ),
    name = "Crime Type"
  ) +
  ggplot2::scale_shape_manual(
    values = c("Total" = 16, "Violent" = 17, "Non-Violent" = 15),
    name = "Crime Type"
  ) +
  ggplot2::labs(
    title = "IRRs Adjusted for Year (Sensitivity Analysis)",
    y     = "Incidence Rate Ratio",
    x     = "Heat Index Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor  = ggplot2::element_blank(),
    legend.position   = "top",
    strip.background  = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text        = ggplot2::element_text(size = 12, face = "bold"),
    axis.title.x      = ggplot2::element_text(size = 14),
    axis.text.x       = ggplot2::element_text(size = 13, angle = 20, hjust = 1),
    axis.title.y      = ggplot2::element_text(size = 14),
    axis.text.y       = ggplot2::element_text(size = 12)
  )

print(p_yr_model)

ggplot2::ggsave(
  filename = "Sensitivity_Year_Adjusted_Model.png",
  plot     = p_yr_model,
  width    = 10,
  height   = 6,
  units    = "in",
  dpi      = 300
)

## Print summary
cat("\n=== SENSITIVITY ANALYSIS: YEAR-ADJUSTED MODELS ===\n")
cat("Models fitted with factor(year) added to primary specification:\n")
cat("n_outcome ~ hi_cat + dow + factor(month) + factor(year)\n\n")
```

## Maximum Heat Index vs Mean Heat Index
```{r}
# Sensitivity Analysis: Using Daily Mean Heat Index Instead of Daily Max

## Calculate Mean HI Percentile Thresholds (85th and 95th)
hi_mean_thresholds <- daily_hi_all %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    p85_mean = quantile(hi_mean_f, 0.85, na.rm = TRUE),
    p95_mean = quantile(hi_mean_f, 0.95, na.rm = TRUE),
    .groups = "drop"
  )

# Display thresholds
hi_mean_thresholds %>%
  dplyr::mutate(
    p85_mean = round(p85_mean, 1),
    p95_mean = round(p95_mean, 1)
  ) %>%
  knitr::kable(
    format = "html",
    col.names = c("City", "85th Percentile (Mean HI °F)", "95th Percentile (Mean HI °F)"),
    caption = "Heat Index Thresholds (Daily Mean HI)"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )

## Create Crime Daily Dataset with Mean HI Categories
crime_daily_meanhi <- crime_all %>%
  dplyr::mutate(
    dow   = lubridate::wday(crime_date, label = TRUE, abbr = TRUE),
    month = lubridate::month(crime_date, label = TRUE, abbr = TRUE),
    year  = lubridate::year(crime_date)
  ) %>%
  dplyr::group_by(city, crime_date, dow, month, year) %>%
  dplyr::summarise(
    n_total      = dplyr::n(),
    n_violent    = sum(toupper(as.character(crime_type)) == "VIOLENT", na.rm = TRUE),
    n_nonviolent = n_total - n_violent,
    .groups = "drop"
  ) %>%
  dplyr::left_join(daily_hi_all, by = c("city", "crime_date" = "date")) %>%
  dplyr::left_join(hi_mean_thresholds, by = "city") %>%
  dplyr::mutate(
    hi_mean_cat = dplyr::case_when(
      is.na(hi_mean_f)                            ~ "Missing",
      hi_mean_f <  p85_mean                       ~ "<85th",
      hi_mean_f >= p85_mean & hi_mean_f < p95_mean ~ "Hot (85–95th)",
      hi_mean_f >= p95_mean                        ~ "Extreme (≥95th)"
    ),
    hi_mean_cat = factor(hi_mean_cat, levels = c("<85th","Hot (85–95th)","Extreme (≥95th)","Missing"))
  )

# Split by city
crime_daily_meanhi_dc  <- crime_daily_meanhi %>% dplyr::filter(city == "DC")
crime_daily_meanhi_bal <- crime_daily_meanhi %>% dplyr::filter(city == "Baltimore")

## Fit Negative Binomial Models Using Mean HI
dc_data_meanhi  <- crime_daily_meanhi_dc  %>% dplyr::filter(hi_mean_cat != "Missing")
bal_data_meanhi <- crime_daily_meanhi_bal %>% dplyr::filter(hi_mean_cat != "Missing")

# DC models
mod_dc_total_meanhi      <- MASS::glm.nb(n_total      ~ hi_mean_cat + dow + factor(month), data = dc_data_meanhi)
mod_dc_violent_meanhi    <- MASS::glm.nb(n_violent    ~ hi_mean_cat + dow + factor(month), data = dc_data_meanhi)
mod_dc_nonviolent_meanhi <- MASS::glm.nb(n_nonviolent ~ hi_mean_cat + dow + factor(month), data = dc_data_meanhi)

# Baltimore models
mod_bal_total_meanhi      <- MASS::glm.nb(n_total      ~ hi_mean_cat + dow + factor(month), data = bal_data_meanhi)
mod_bal_violent_meanhi    <- MASS::glm.nb(n_violent    ~ hi_mean_cat + dow + factor(month), data = bal_data_meanhi)
mod_bal_nonviolent_meanhi <- MASS::glm.nb(n_nonviolent ~ hi_mean_cat + dow + factor(month), data = bal_data_meanhi)

## Extract IRRs for Mean HI Models (reuses your irr_df_custom() helper)
irr_dc_meanhi <- dplyr::bind_rows(
  irr_df_custom(mod_dc_total_meanhi,      "Total",       "DC", "hi_mean_cat"),
  irr_df_custom(mod_dc_violent_meanhi,    "Violent",     "DC", "hi_mean_cat"),
  irr_df_custom(mod_dc_nonviolent_meanhi, "Non-Violent", "DC", "hi_mean_cat")
)

irr_bal_meanhi <- dplyr::bind_rows(
  irr_df_custom(mod_bal_total_meanhi,      "Total",       "Baltimore", "hi_mean_cat"),
  irr_df_custom(mod_bal_violent_meanhi,    "Violent",     "Baltimore", "hi_mean_cat"),
  irr_df_custom(mod_bal_nonviolent_meanhi, "Non-Violent", "Baltimore", "hi_mean_cat")
)

# Combine and standardize labels (same pattern as irr_all_temp)
irr_all_meanhi <- dplyr::bind_rows(irr_dc_meanhi, irr_bal_meanhi) %>%
  dplyr::mutate(
    city = dplyr::recode(city,
      "Baltimore" = "Baltimore, MD",
      "DC"        = "DC"
    ),
    hi_level = stringr::str_squish(hi_level),
    hi_level = dplyr::case_when(
      stringr::str_detect(hi_level, "Hot")     ~ "Hot (85–95th)",
      stringr::str_detect(hi_level, "Extreme") ~ "Extreme (≥95th)",
      TRUE ~ hi_level
    )
  ) %>%
  dplyr::rename(IRR = irr, LCL = lower.CL, UCL = upper.CL)

## Compare Max HI vs Mean HI Models
comparison_maxhi_vs_meanhi <- irr_all_three %>%
  dplyr::select(city, outcome, hi_level, IRR_Max = IRR, LCL_Max = LCL, UCL_Max = UCL) %>%
  dplyr::left_join(
    irr_all_meanhi %>% dplyr::select(city, outcome, hi_level, IRR_Mean = IRR, LCL_Mean = LCL, UCL_Mean = UCL),
    by = c("city", "outcome", "hi_level")
  ) %>%
  dplyr::mutate(
    Max_HI_Model  = sprintf("%.3f (%.3f–%.3f)", IRR_Max,  LCL_Max,  UCL_Max),
    Mean_HI_Model = sprintf("%.3f (%.3f–%.3f)", IRR_Mean, LCL_Mean, UCL_Mean)
  )

# Display comparison table
comparison_maxhi_vs_meanhi %>%
  dplyr::select(City = city, `Crime Type` = outcome, `Heat Category` = hi_level,
                `Max HI Model IRR (95% CI)`  = Max_HI_Model,
                `Mean HI Model IRR (95% CI)` = Mean_HI_Model) %>%
  kable(
    format = "html",
    caption = "Sensitivity Analysis: Comparison of Max Heat Index vs Mean Heat Index Models"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) %>%
  collapse_rows(columns = 1:2, valign = "top")

# Save comparison
readr::write_csv(comparison_maxhi_vs_meanhi, "sensitivity_analysis_MaxHI_vs_MeanHI.csv")

## Visualize Mean HI Model Results
y_min_meanhi <- min(irr_all_meanhi$LCL, na.rm = TRUE) * 0.95
y_max_meanhi <- max(irr_all_meanhi$UCL, na.rm = TRUE) * 1.05

p_meanhi_model <- ggplot2::ggplot(
  irr_all_meanhi,
  ggplot2::aes(
    x = hi_level, y = IRR, ymin = LCL, ymax = UCL,
    shape = outcome, color = outcome
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, color = "grey50") +
  ggplot2::geom_pointrange(
    position = ggplot2::position_dodge(width = 0.45),
    size = 1.2,
    fatten = 7
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_y_log10(limits = c(y_min_meanhi, y_max_meanhi)) +
  ggplot2::scale_color_manual(
    values = c(
      "Total"       = "#440154",
      "Violent"     = "#31688e",
      "Non-Violent" = "#35b779"
    ),
    name = "Crime Type"
  ) +
  ggplot2::scale_shape_manual(
    values = c("Total" = 16, "Violent" = 17, "Non-Violent" = 15),
    name = "Crime Type"
  ) +
  ggplot2::labs(
    title = "IRRs Using Daily Mean Heat Index (Sensitivity Analysis)",
    y     = "Incidence Rate Ratio",
    x     = "Mean Heat Index Category (vs <85th percentile)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor  = ggplot2::element_blank(),
    legend.position   = "top",
    strip.background  = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text        = ggplot2::element_text(size = 12, face = "bold"),
    axis.title.x      = ggplot2::element_text(size = 14),
    axis.text.x       = ggplot2::element_text(size = 13, angle = 20, hjust = 1),
    axis.title.y      = ggplot2::element_text(size = 14),
    axis.text.y       = ggplot2::element_text(size = 12)
  )

print(p_meanhi_model)

ggplot2::ggsave(
  filename = "Sensitivity_MeanHI_Model.png",
  plot     = p_meanhi_model,
  width    = 10,
  height   = 6,
  units    = "in",
  dpi      = 300
)

## Correlation check: how closely do Max HI and Mean HI track each other?
hi_corr <- daily_hi_all %>%
  dplyr::group_by(city) %>%
  dplyr::summarise(
    correlation = cor(hi_max_f, hi_mean_f, use = "complete.obs"),
    .groups = "drop"
  )
print(hi_corr)

## Print summary
cat("\n=== SENSITIVITY ANALYSIS: MAX HI VS MEAN HI ===\n")
cat("Models refit using daily MEAN heat index category in place of daily MAX:\n")
cat("n_outcome ~ hi_mean_cat + dow + factor(month)\n\n")
cat("Correlation between max and mean HI by city:\n")
print(hi_corr)
```


Look into continuous splines and IRR overlayed
(Comparison of the two heat categorization methods)

```{r}
## Overlay: Categorical Model Predictions on Spline Dose-Response Curve
# Get predicted counts (not IRR) from the categorical models, at the SAME
# reference level (Wed, July) used in the spline predictions, so both are
# directly comparable on the response (count) scale.

get_cat_predictions <- function(model, data, city_label, outcome_label) {

  emm <- emmeans::emmeans(
    model, ~ hi_cat,
    at = list(dow = "Wed", month = "Jul"),
    type = "response"
  )

  emm_df <- as.data.frame(emm)

  # Harmonize column names across emmeans output versions
  nm <- names(emm_df)
  if (!"response" %in% nm) {
    cand <- intersect(c("rate", "prob", "mu"), nm)
    if (length(cand) > 0) names(emm_df)[names(emm_df) == cand[1]] <- "response"
  }
  if (!"asymp.LCL" %in% nm & "lower.CL" %in% nm) names(emm_df)[names(emm_df) == "lower.CL"] <- "asymp.LCL"
  if (!"asymp.UCL" %in% nm & "upper.CL" %in% nm) names(emm_df)[names(emm_df) == "upper.CL"] <- "asymp.UCL"

  emm_df %>%
    dplyr::transmute(
      hi_cat   = hi_cat,
      pred     = response,
      lower    = asymp.LCL,
      upper    = asymp.UCL,
      city     = city_label,
      outcome  = outcome_label
    )
}

cat_pred_all <- dplyr::bind_rows(
  get_cat_predictions(mod_dc_violent,     dc_data,  "DC",        "Violent"),
  get_cat_predictions(mod_dc_nonviolent,  dc_data,  "DC",        "Non-Violent"),
  get_cat_predictions(mod_bal_violent,    bal_data, "Baltimore", "Violent"),
  get_cat_predictions(mod_bal_nonviolent, bal_data, "Baltimore", "Non-Violent")
)

## Match city labels to spline figure
cat_pred_all$city <- dplyr::recode(cat_pred_all$city,
  "DC"        = "Washington, DC",
  "Baltimore" = "Baltimore, MD"
)

## Attach x-axis segment ranges for each category, using each city's own
# observed HI range and percentile thresholds
cat_pred_all <- cat_pred_all %>%
  dplyr::left_join(
    thresh_lines %>% dplyr::select(city, p85, p95),
    by = "city"
  ) %>%
  dplyr::left_join(
    pred_all_ns %>%
      dplyr::group_by(city) %>%
      dplyr::summarise(hi_min = min(hi_max_f), hi_max_obs = max(hi_max_f), .groups = "drop"),
    by = "city"
  ) %>%
  dplyr::mutate(
    x_start = dplyr::case_when(
      hi_cat == "<85th"                ~ hi_min,
      hi_cat == "Hot (85–95th)"        ~ p85,
      hi_cat == "Extreme (≥95th)"      ~ p95
    ),
    x_end = dplyr::case_when(
      hi_cat == "<85th"                ~ p85,
      hi_cat == "Hot (85–95th)"        ~ p95,
      hi_cat == "Extreme (≥95th)"      ~ hi_max_obs
    ),
    outcome = factor(outcome, levels = c("Violent", "Non-Violent")),
    city    = factor(city, levels = c("Washington, DC", "Baltimore, MD"))
  )

## Plot: spline curve + categorical step function overlay
p_overlay <- ggplot2::ggplot() +
  # Spline curve + ribbon (same as before)
  ggplot2::geom_ribbon(
    data = pred_all_ns,
    ggplot2::aes(x = hi_max_f, ymin = lower, ymax = upper, fill = outcome),
    alpha = 0.12, color = NA
  ) +
  ggplot2::geom_line(
    data = pred_all_ns,
    ggplot2::aes(x = hi_max_f, y = fit, color = outcome),
    linewidth = 1.1
  ) +
  # Percentile threshold lines
  ggplot2::geom_vline(
    data = thresh_lines, ggplot2::aes(xintercept = p85),
    linetype = "dashed", color = "grey40", linewidth = 0.6
  ) +
  ggplot2::geom_vline(
    data = thresh_lines, ggplot2::aes(xintercept = p95),
    linetype = "dotted", color = "grey20", linewidth = 0.6
  ) +
  # Categorical model's step-function prediction, as horizontal segments
  ggplot2::geom_segment(
    data = cat_pred_all,
    ggplot2::aes(x = x_start, xend = x_end, y = pred, yend = pred, color = outcome),
    linewidth = 1.4, linetype = "solid", alpha = 0.9
  ) +
  # CI for each categorical step, as a shaded rectangle
  ggplot2::geom_rect(
    data = cat_pred_all,
    ggplot2::aes(xmin = x_start, xmax = x_end, ymin = lower, ymax = upper, fill = outcome),
    alpha = 0.15, inherit.aes = FALSE
  ) +
  ggplot2::facet_wrap(~ city, ncol = 2) +
  ggplot2::scale_color_manual(
    values = c("Violent" = "#31688e", "Non-Violent" = "#35b779"),
    name = "Crime Type"
  ) +
  ggplot2::scale_fill_manual(
    values = c("Violent" = "#31688e", "Non-Violent" = "#35b779"),
    name = "Crime Type"
  ) +
  ggplot2::labs(
    title = "Spline (Continuous) vs. Categorical Model Predictions",
    subtitle = "Thin line + shaded ribbon = spline; thick flat segments + boxes = categorical model (<85th / Hot / Extreme)",
    y = "Predicted Daily Crime Count",
    x = "Daily Maximum Heat Index (°F)"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "top",
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.text = ggplot2::element_text(size = 12, face = "bold")
  )

print(p_overlay)

ggplot2::ggsave(
  filename = "Overlay_Spline_vs_Categorical.png",
  plot     = p_overlay,
  width    = 10,
  height   = 7,
  units    = "in",
  dpi      = 300
)
```

```{r}
## Get peak HI value and predicted count for each city/outcome combination
peak_summary <- pred_all_ns %>%
  dplyr::group_by(city, outcome) %>%
  dplyr::filter(fit == max(fit)) %>%
  dplyr::select(city, outcome, hi_max_f, fit) %>%
  dplyr::ungroup()

print(peak_summary)
```



